---
title: "stubthat"
author: "Sainath Adapa"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{stubthat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Introduction
stubthat package provides stubs meant to be used while unit testing in R. The API is highly inspired by the [Sinon.js](http://sinonjs.org/). This package is meant to be used along with the [testthat](https://cran.r-project.org/web/packages/testthat/index.html) package, specifically the 'with_mock' function.

```{r include=FALSE}
library('stubthat')
```

# Usage
There are mainly three steps to the creating a stub - 
1. Invoke the stub function with the function that needs to be mocked
```{r}
jedi_or_sith <- function(x) return('No one')
stub_builder <- stub(jedi_or_sith)
```
2. Define the behavior. This is explained in detail in the next section.
```{r}
stub_builder$withArgs(x = 'Luke')$returns('Jedi')
```
3. Build the stub.
```{r results='asis'}
stub_of_fun <- stub_builder$build()
```
Once the stub is built, you can use the returned function.
```{r}
jedi_or_sith('Luke')
stub_of_fun('Luke')
```

# Behaviors
- ````stub_builder$returns(10)````
- ````stub_builder$throws('err msg')````
- ````stub_builder$expects(a = 1, b = 2, d = 3, c = 4)````
- ````stub_builder$withExactArgs(a = 1, b = 2, d = 3, c = 4)$returns(10)````
- ````stub_builder$withExactArgs(a = 1, b = 2, d = 3, c = 4)$throws('err_msg')````
- ````stub_builder$withArgs(a = 1, d = 3)$returns(10)````
- ````stub_builder$withArgs(a = 1, d = 3)$throws('err msg')````
- ````stub_builder$onCall(3)$withExactArgs(a = 1, b = 2, d = 3, c = 4)$returns(10)````
- ````stub_builder$onCall(2)$withExactArgs(a = 1, b = 2, d = 3, c = 4)$throws('err_msg')````
- ````stub_builder$onCall(2)$returns(10)````
- ````stub_builder$onCall(2)$throws('err_msg')````
- ````stub_builder$onCall(3)$withArgs(a = 1, d = 3)$returns(10)````
- ````stub_builder$onCall(3)$withArgs(a = 1, d = 3)$throws('err msg')````
