---
title: "stubthat - Stubs for unit testing in R"
author: "Sainath Adapa"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
    toc: true
vignette: >
  %\VignetteIndexEntry{stubthat - Stubs for unit testing in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
library(stubthat)
```

# Introduction
*stubthat* package provides stubs for use while unit testing in R. The API is highly inspired by the *[Sinon.js](http://sinonjs.org/)*. This package is meant to be used along with the *[testthat](https://cran.r-project.org/package=testthat)* package, specifically the *'with_mock'* function.

To understand what a stub is and how they are used while unit testing, please take a look at this Stackoverflow question [What is a “Stub”?](http://stackoverflow.com/questions/463278/what-is-a-stub).


# Usage
There are three main steps for creating & using a stub of a function -  

- Invoke the *stub* function with the function that needs to be mocked
```{r}
jedi_or_sith <- function(x) return('No one')
jedi_or_sith_stub <- stub(jedi_or_sith)
```
- Define the behavior. This is explained in detail in the API section.
```{r}
jedi_or_sith_stub$withArgs(x = 'Luke')$returns('Jedi')
```
- Once the behavior is defined, you can use the stub by calling the `jedi_or_sith_stub$f` function.
```{r}
jedi_or_sith('Luke')
jedi_or_sith_stub$f('Luke')
```

# Use cases
Stubs are generally used for testing purposes. Here is an example:
```{r}
library(httr) # provides the GET and status_code functions

url_downloader <- function(url) GET(url)

check_api_endpoint_status <- function(url) {
  response <- url_downloader(url)
  response_status <- status_code(response)
  ifelse(response_status == 200, 'up', 'down')
}
```
This function *check_api_endpoint_status* should make a *GET* request (via the url_downloader function) to the specified url (say https://example.com/endpoint) and it should return *'up'* if the status code is *'200'*. Return *'down'* otherwise. While testing, it is generally a good idea to avoid making repeated (or any) requests to external sources.

Using stubs (and `with_mock` from [mockr](https://github.com/krlmlr/mockr)), the above function can be tested without accessing the external source, as shown below:
```{r}
url_downloader_stub <- stub(url_downloader)
url_downloader_stub$withArgs(url = 'good url')$returns(200)
url_downloader_stub$withArgs(url = 'bad url')$returns(404)

library('testthat') # provides the expect_equal function
library('mockr') # provides the with_mock function

check_api_endpoint_status_tester <- function(x) {
  with_mock(url_downloader = url_downloader_stub$f,
            check_api_endpoint_status(x))
}

expect_equal(check_api_endpoint_status_tester('good url'), 'up')
expect_equal(check_api_endpoint_status_tester('bad url'),  'down')
```


# API

## Check if the stub is called with specified arguments

### `stub$expects(...)`
Stub will check the incoming arguments for the specified set of arguments. Throws an error if there is a mismatch.
```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$expects(a = 2)

stub_of_sum$f(2)
stub_of_sum$f(3)
```

### `stub$strictlyExpects(...)`
The set of specified arguments should be exactly matched with the set of incoming arguments.
```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$strictlyExpects(a = 2)
stub_of_sum$f(2)
```
The above call resulted in the error because the incoming set of arguments was `a = 2, b = 1`, but the defined set of expected arguments consisted only `a = 2`.

```{r error=TRUE}
stub_of_sum$strictlyExpects(a = 2, b = 1)
stub_of_sum$f(2)
```

### `stub$onCall(#)$expects(...)`

The stub expects the specifed arguments on the *nth* call.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$onCall(3)$expects(a = 2)

stub_of_sum$f(100)
stub_of_sum$f(100)
stub_of_sum$f(100)
```

### `stub$onCall(#)$strictlyExpects(...)`

The stub expects the **exact** set of specifed arguments on the *nth* call.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$onCall(3)$strictlyExpects(a = 2, b = 2)

stub_of_sum$f(2)
stub_of_sum$f(2)
stub_of_sum$f(2)
```

## Make the stub return a specified value

### `stub$returns(...)`

Unless otherwise specified, the stub always returns the specified value.
```{r}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$returns(0)

stub_of_sum$f(2)
```

### `stub$onCall(#)$returns(...)`

The stub returns the specified value on the *nth* call.

```{r}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$onCall(2)$returns(0)

stub_of_sum$f(2)
stub_of_sum$f(2)
```

### `stub$withArgs(...)$returns(...)`

The stub returns the specified value when it is called with the specified arguments.

```{r}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$withArgs(a = 2)$returns(0)

stub_of_sum$f(1)
stub_of_sum$f(2)
```

### `stub$withExactArgs(...)$returns(...)`

The stub returns the specified value when it is called with the **exact** set of specified arguments.

```{r}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$withExactArgs(a = 2)$returns(0) # won't work because value for b is not defined
stub_of_sum$withExactArgs(a = 2, b = 1)$returns(1)

stub_of_sum$f(1)
stub_of_sum$f(2)
```


## Make the stub throw an error with a specified message

### `stub$throws('')`

Unless otherwise specified, the stub throws an error with the specified message.
```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$throws('some err msg')

stub_of_sum$f(2)
```

### `stub$onCall(#)$throws('')`

The stub throws an error on the *nth* call.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$onCall(2)$throws('some err msg')

stub_of_sum$f(0)
stub_of_sum$f(0)
```

### `stub$withArgs(...)$throws('')`

The stub throws an error when it is called with the specified arguments.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$withArgs(a = 2)$throws('some err msg')

stub_of_sum$f(1)
stub_of_sum$f(2)
```

### `stub$withExactArgs(...)$throws('')`

The stub returns the specified value when it is called with the **exact** set of specified arguments.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$withExactArgs(a = 2)$throws('good') # won't work because value for b is not defined
stub_of_sum$withExactArgs(a = 2, b = 1)$throws('nice')

stub_of_sum$f(1)
stub_of_sum$f(2)
```



## Get the number of times the stub has been called

### `stub$calledTimes()`

Using this, one can obtain the number of times, the stub has been called.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

ans <- stub_of_sum$f(3)
ans <- stub_of_sum$f(3)
stub_of_sum$calledTimes()
ans <- stub_of_sum$f(3)
stub_of_sum$calledTimes()
```

## Extra

Convenience functions to reduce repetition of code.

### `stub$onCall(#)$expects(...)$returns(...)`

On *nth* call, the stub will check for the specified arguments, and if satisfied, returns the specified value.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$onCall(1)$expects(a = 1)$returns('good')
stub_of_sum$onCall(3)$expects(a = 3)$returns('nice')

stub_of_sum$f(3)
stub_of_sum$f(3)
stub_of_sum$f(3)
```
This is same as calling `stub$onCall(#)$expects(...)` and `stub$onCall(#)$returns(...)` separately.
```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$onCall(1)$expects(a = 1)
stub_of_sum$onCall(1)$returns('good')
stub_of_sum$onCall(3)$returns('nice')
stub_of_sum$onCall(3)$expects(a = 3)

stub_of_sum$f(3)
stub_of_sum$f(3)
stub_of_sum$f(3)
```

### `stub$onCall(#)$strictlyExpects(...)$returns(...)`

On *nth* call, the stub will check for the **exact** set of specified arguments, and if satisfied, returns the specified value.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$onCall(1)$strictlyExpects(a = 3)$returns('good')
stub_of_sum$onCall(3)$strictlyExpects(a = 3, b = 1)$returns('nice')

stub_of_sum$f(3)
stub_of_sum$f(3)
stub_of_sum$f(3)
```

### `stub$onCall(#)$expects(...)$throws('')`

On *nth* call, the stub will check for the specified arguments, and if satisfied, throws an error with the specified message.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$onCall(1)$expects(a = 1)$throws('good')
stub_of_sum$onCall(3)$expects(a = 3)$throws('nice')

stub_of_sum$f(3)
stub_of_sum$f(3)
stub_of_sum$f(3)
```

### `stub$onCall(#)$strictlyExpects(...)$throws('')`

On *nth* call, the stub will check for the **exact** set of specified arguments, and if satisfied, throws an error with the specified message.

```{r error=TRUE}
sum <- function(a, b = 1) return(a + b)
stub_of_sum <- stub(sum)

stub_of_sum$onCall(1)$strictlyExpects(a = 3)$throws('good')
stub_of_sum$onCall(3)$strictlyExpects(a = 3, b = 1)$throws('nice')

stub_of_sum$f(3)
stub_of_sum$f(3)
stub_of_sum$f(3)
```

